---
title: "Sumo Kimarite"
output: html_notebook
---

Functions to interface with [Sumo Reference](http://sumodb.sumogames.de/):
```{r}
source("sumodb.R")
```

&nbsp;  
&nbsp;  
&nbsp;  
[Makuuchi](https://en.wikipedia.org/wiki/Makuuchi) (top division) results in 2016:
```{r}
bout <- sumodbBoutQuery(basho = "2016", division = "m")
```

&nbsp;  
&nbsp;  
&nbsp;  
Most frequent [kimarite](https://en.wikipedia.org/wiki/Kimarite) is `r sprintf("%s", (bout %>% count(kimarite) %>% arrange(-n))[1, 1])` --- `r sprintf("%.1f%%", (bout %>% count(kimarite) %>% arrange(-n))[1, 2] / nrow(bout) * 100)` of all bouts (including [fusen](https://en.wiktionary.org/wiki/fusen)):
```{r}
bout %>% count(kimarite) %>% arrange(-n)
```

&nbsp;  
&nbsp;  
&nbsp;  
[Banzuke](https://en.wikipedia.org/wiki/Banzuke) for each tournament:
```{r}
banzuke <- do.call(
	rbind,
	lapply(
		c("01", "03", "05", "07", "09", "11"),
		function(x) sumodbBanzukeQuery(paste0("2016", x)) %>% mutate(basho = paste("2016", x, sep = "."))
	)
) %>%
	setNames(tolower(names(.))) %>%
	select(
		basho,
		rikishi,
		`height/weight`
	) %>%
	mutate(
		height = as.numeric(str_match(`height/weight`, "([0-9.]+) cm")[, 2]),
		weight = as.numeric(str_match(`height/weight`, "([0-9.]+) kg")[, 2])
	) %>%
	select(-`height/weight`)
banzuke
```

&nbsp;  
&nbsp;  
&nbsp;  
Add height/weight data to bout results:
```{r}
suppressMessages(
	bout <- left_join(
		left_join(
			bout %>% mutate(rikishi = ifelse(win1 == 1, shikona1, shikona2)),
			banzuke
		) %>%
			rename(winner.height = height, winner.weight = weight) %>%
			mutate(rikishi = ifelse(win1 == 0, shikona1, shikona2)),
		banzuke
	) %>%
		rename(loser.height = height, loser.weight = weight)
)
bout
```

&nbsp;  
&nbsp;  
&nbsp;  
Bundle less frequent kimarite as *other*:
```{r}
suppressMessages(
	bout <- inner_join(
		bout,
		bout %>%
			count(kimarite) %>%
			arrange(-n) %>%
			mutate(
				freq_rank = row_number(),
				kimarite2 = ifelse(freq_rank > 7, "other", kimarite)
			)
	)
)
bout
```


```{r}
suppressPackageStartupMessages(
	library(plotly)
)
  
bout %>%
	plot_ly() %>%
	add_markers(
		x = ~winner.weight,
		y = ~loser.weight,
		color = ~kimarite2
	)
```
