---
title: "Distribution of kimarite (winning techniques) by weight and height of rikishi (sumo wrestlers)"
output: html_notebook
---

```{r, message=FALSE, warning=FALSE}
library(dplyr)
library(plotly)
options(stringsAsFactors = FALSE)
```

&nbsp;  
&nbsp;  
&nbsp;  
Three years, `r 3 * 6` tournaments:
```{r}
tournament <- apply(
	expand.grid(
		2014:2016, # years
		sprintf("%02d", seq(1, 11, 2)) # six tournaments a year
	),
	1,
	paste, collapse = "."
)
  
m <- matrix(tournament[order(tournament)], ncol = 6, byrow = TRUE)
  
prmatrix(
	m,
	rowlab = rep("", ncol(m)),
	collab = rep("", ncol(m)),
	quote = FALSE
)

```

&nbsp;  
&nbsp;  
&nbsp;  
Fetch [makuuchi](https://en.wikipedia.org/wiki/Makuuchi) (top division) results from [Cervus1983/sumodb](https://github.com/Cervus1983/sumodb) repository:
```{r}
bout <- do.call(
	rbind,
	lapply(
		tournament,
		function(x) read.csv(
			paste0("https://raw.githubusercontent.com/Cervus1983/sumodb/master/CSV/", x, ".results.csv")
		)
	)
)
  
bout
```

&nbsp;  
&nbsp;  
&nbsp;  
Most frequent [kimarite](https://en.wikipedia.org/wiki/Kimarite) is `r sprintf("%s", (bout %>% count(kimarite) %>% arrange(-n))[1, 1])` --- `r sprintf("%.1f%%", (bout %>% count(kimarite) %>% arrange(-n))[1, 2] / nrow(bout) * 100)` of all bouts (including forfeits --- [fusen](https://en.wiktionary.org/wiki/fusen)):
```{r}
bout %>% count(kimarite) %>% arrange(-n)
```

&nbsp;  
&nbsp;  
&nbsp;  
[Banzuke](https://en.wikipedia.org/wiki/Banzuke) for each tournament:
```{r}
banzuke <- do.call(
	rbind,
	lapply(
		tournament,
		function(x) read.csv(
			paste0("https://raw.githubusercontent.com/Cervus1983/sumodb/master/CSV/", x, ".banzuke.csv")
		)
	)
)
  
banzuke
```

&nbsp;  
&nbsp;  
&nbsp;  
Height & weight distribution:
```{r}
banzuke %>%
	group_by(rikishi) %>%
	summarise(height = mean(height)) %>%
	plot_ly(width = 910) %>%
	add_histogram(
		x = ~height,
		xbins = list(
			start = floor(min(banzuke$height)),
			end = ceiling(max(banzuke$height)),
			size = 1
		)
	) %>%
	layout(
		xaxis = list(title = "Height (cm)"),
		yaxis = list(
			showgrid = FALSE,
			showticklabels = FALSE
		)
	)

banzuke %>%
	group_by(rikishi) %>%
	summarise(weight = mean(weight)) %>%
	plot_ly(width = 910) %>%
	add_histogram(
		x = ~weight,
		xbins = list(
			start = floor(min(banzuke$weight)),
			end = ceiling(max(banzuke$weight)),
			size = 5
		)
	) %>%
	layout(
		xaxis = list(title = "Weight (kg)"),
		yaxis = list(
			showgrid = FALSE,
			showticklabels = FALSE
		)
	)
```

&nbsp;  
&nbsp;  
&nbsp;  
Add height & weight data to bout results, group less frequent kimarite as *other*, and plot kimarite distribution:
```{r, message=FALSE}
bout2 <- left_join(
	left_join(
		bout %>% mutate(rikishi = ifelse(win1 == 1, shikona1, shikona2)),
		banzuke
	) %>%
		select(-rank) %>%
		rename(winner.height = height, winner.weight = weight) %>%
		mutate(rikishi = ifelse(win1 == 0, shikona1, shikona2)),
	banzuke
) %>%
	select(-rank) %>%
	rename(loser.height = height, loser.weight = weight)
  
bout3 <- inner_join(
	bout2,
	bout2 %>%
		count(kimarite) %>%
		arrange(-n) %>%
		mutate(
			freq_rank = row_number(),
			kimarite2 = ifelse(freq_rank > 7, "other", kimarite) # top 7 kimarite, everything else as "other"
		) %>%
		select(kimarite, kimarite2)
) %>% filter(
		complete.cases(.), # omit records with missing weight/height
		kimarite != "fusen" # omit forfeits
	)
  
bout3 %>%
	plot_ly(width = 910) %>%
	add_markers(
		x = ~winner.height,
		y = ~loser.height,
		color = ~kimarite2,
		hoverinfo = "text",
		text = ~paste(basho, "~", shikona1, "v", shikona2, "~", kimarite)
	) %>%
	layout(
		xaxis = list(
			title = "Winner's height"
		),
		yaxis = list(
			title = "Loser's height"
		)
	)

bout3 %>%
	plot_ly(width = 910) %>%
	add_markers(
		x = ~winner.weight,
		y = ~loser.weight,
		color = ~kimarite2,
		hoverinfo = "text",
		text = ~paste(basho, "~", shikona1, "v", shikona2, "~", kimarite)
	) %>%
	layout(
		xaxis = list(
			title = "Winner's weight"
		),
		yaxis = list(
			title = "Loser's weight"
		)
	)
```

&nbsp;  